---
title: "Who Is Who?"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(kabrutils)
library(DBI)
library(testdat)
library(shinyvalidate)

load("data/candidates-authors.Rdata")

db_tbl_er_candidates <- "er_candidates"
```


```{r}
# Update

# - get new additional_infos in create-input-data.R
# - add new source to get_candidates_names()
```

```{r}
#' get a candidate pair to check
#' 
#' rule to choose
#' - not yet checked: `decision == "no_judgement"`
#' - minimum rank: highest similarity
#' - minimum id: in order to choose when similar rank
#' 
#' @param data dataframe with similarties
#' @return dataframe with 1 row
#' 
get_candidates_to_check <- function(data) {
  data %>% 
    dplyr::filter(decision == "no_judgement" & rank == min(rank), row_number() == 1)
}


#' Get candidates names
#' 
#' Looks for id_1 and if source is books or poster. 
#' Then joins name from additional_infos df depending in source
#' Does same for id_2.
#' Then joins datas for both together
#' 
#' not really good implemented yet,  because dependent on candidates$books_additional_infos which is quite unstructured
#' 
get_candidates_names <- function(data) {
  candidate_1 <- data %>% 
    dplyr::distinct(id, source_1, id_1)
  print(candidate_1)
  
  if(candidate_1$source_1 == "book_author") {
    candidate_1 <- candidate_1 %>% 
      dplyr::left_join(candidates$books_additional_infos %>% 
                         dplyr::distinct(author_id, name_1 = author), by = c("id_1" = "author_id"))
  } else if (candidate_1$source_1 == "poster_author"){ 
    candidate_1 <- candidate_1 %>% 
      dplyr::left_join(candidates$poster_additional_infos %>% 
                         dplyr::distinct(author_id, name_1 = author), by = c("id_1" = "author_id"))
    
  } else if (candidate_1$source_1 == "book_publisher"){ 
    print("here")
    candidate_1 <- candidate_1 %>% 
      dplyr::left_join(candidates$books_publisher_additional_infos %>% 
                         dplyr::distinct(publisher_id, name_1 = publisher), by = c("id_1" = "publisher_id"))
  } else {
    print("not implemented")
  }
  
  candidate_2 <- data %>% 
    dplyr::distinct(id, source_2, id_2)
  if(candidate_2$source_2 == "book_author") {
    candidate_2 <- candidate_2 %>% 
      dplyr::left_join(candidates$books_additional_infos %>% 
                         dplyr::distinct(author_id, name_2 = author), by = c("id_2" = "author_id"))
  } else if (candidate_2$source_2 == "poster_author"){ 
    candidate_2 <- candidate_2 %>% 
      dplyr::left_join(candidates$poster_additional_infos %>% 
                         dplyr::distinct(author_id, name_2 = author), by = c("id_2" = "author_id"))
    
  } else if (candidate_2$source_2 == "book_publisher"){ 
    print("here")
    candidate_2 <- candidate_2 %>% 
      dplyr::left_join(candidates$books_publisher_additional_infos %>% 
                         dplyr::distinct(publisher_id, name_2 = publisher), by = c("id_2" = "publisher_id"))
  } else {
    print("not implemented")
  }
  
  candidate_names <- dplyr::inner_join(candidate_1, candidate_2, by = "id")
  candidate_names
}


#' Save the decision in dataframe and as a csv file
#' 
#' Includes test after saving, that decision was saved in DB.
#'
#' @param data data with similarities
#' @param candidates the two candidates to decide on
#' @param decision: positive, negatove, not_sure, next (no decision)
#' 
save_decision <- function(candidates, decision, ...) {
  
  this_candidate_id <- candidates$id
  this_decision <- decision

  if (decision == "positive") {
    if (input$keep_candidate_2_label == TRUE) {
      query <- paste0("UPDATE ", db_tbl_er_candidates, " SET decision = '", this_decision, "', keep_label_2 = 1 WHERE id =", this_candidate_id, ";")
    } else {
      query <- paste0("UPDATE ", db_tbl_er_candidates, " SET decision = '", this_decision, "' WHERE id =", this_candidate_id, ";")
    }

    con <- connect_db()
    dbExecute(con, query)
    test_that(
      desc = "double-check if this_decision was saved in db",
      expect_equal(
        tbl(con, db_tbl_er_candidates) %>%
          filter(id == this_candidate_id, decision == this_decision) %>%
          collect() %>% nrow(),
        1
      )
    )

    dbDisconnect(con)
    rm(con)

    print("it's positive, baby")
  } else if (decision == "negative") {
    query <- paste0("UPDATE ", db_tbl_er_candidates, " SET decision = '", this_decision, "' WHERE id =", this_candidate_id, ";")
    con <- connect_db()
    dbExecute(con, query)
    test_that(
      desc = "double-check if this_decision was saved in db",
      expect_equal(
        tbl(con, db_tbl_er_candidates) %>%
          filter(id == this_candidate_id, decision == this_decision) %>%
          collect() %>% nrow(),
        1
      )
    )

    dbDisconnect(con)
    rm(con)
    print("nope, not the same")
  } else if (decision == "not_sure") {
    query <- paste0("UPDATE ", db_tbl_er_candidates, " SET decision = '", this_decision, "' WHERE id =", this_candidate_id, ";")
    con <- connect_db()
    dbExecute(con, query)

    test_that(
      desc = "double-check if this_decision was saved in db",
      expect_equal(
        tbl(con, db_tbl_er_candidates) %>%
          filter(id == this_candidate_id, decision == this_decision) %>%
          collect() %>% nrow(),
        1
      )
    )

    dbDisconnect(con)
    rm(con)
    print("can't decide, sorry. needs to be checked again")
  } else if (decision == "next_candidates") {
    print("want to check next candidates")
    return(data)
  }
}

#' Get entities that are available for resultion
#' 
#' Basically a distinct on db_tbl_er_candidates entities col.
#' Is then used to fill a dropdown.
#' @return df with distinct(er_candidates, entities)

get_er_entities <- function() {
  con <- connect_db()
  entities <- tbl(con, db_tbl_er_candidates) %>% 
    distinct(entities) %>% 
    collect()
  dbDisconnect(con)
  rm(con)
  return(entities)
}

```


Candidates ER
=====================================  

Beschreibung {.sidebar}
-------------------------------------

Welche Entitäten sollen aufgelöst werden?

```{r}
entities_list <- get_er_entities() %>% 
  filter(entities != "manual") %>% 
  as.list()
names(entities_list) <- entities_list
selectInput("entities", label = NULL, choices = entities_list)
```

Bei `Start/Next` wird das nächste Kandidatenpaar angezeigt.

```{r}
actionButton("button_start", "Start/Next", class = "btn-primary")
```


### Anmerkungen

* Wenn du dich für `Ja, identisch` entscheidest, wird der Name/das Label von `Candidate 1` behalten. Möchtest du das ändern und lieber den Namen von `Candidate 2` behalten, dann setze bitte den Hacken über dem `Ja`-Button.

* Die Information wird direkt in der Datenbank gespeichert, aber der Merge an sich wird manuell angestoßen.

Column {data-width=30%}
-------------------------------------

### Entscheidung

Sind die beiden Entitäten identisch oder nicht? Bist du unsicher?

```{r show_buttons}
checkboxInput("keep_candidate_2_label", "Behalte Name von Candidate 2")
actionButton("button_positive", "Ja, identisch", class = "btn-primary")
hr(); br()
actionButton("button_negative", "Nein, verschieden", class = "btn-primary")
br();hr(); br()
actionButton("button_not_sure", "Unsicher", class = "btn-primary")
br();br()
```

```{r get_candidates_from_db}

candidates_to_check <- eventReactive(input$button_start, {
   these_entities <- input$entities
   print(these_entities)
    con <- connect_db()
    candidates <- tbl(con, db_tbl_er_candidates) %>% 
      dplyr::filter(entities == these_entities, decision == "no_judgement") %>% 
      dplyr::arrange(rank) %>% 
      dplyr::filter(row_number() == 1) %>% 
      dplyr::collect()
    
    DBI::dbDisconnect(con); rm(con)
    return(candidates)
  })

```

### New Label

Are both provided labels not correct? Then insert a new one:

```{r new_name}

# it's possible that both names are not correct
# here you can set a new name

textInput("new_label", "Insert New Label", NA)
actionButton("new_label_submit", "Submit")

new_label <- reactive({
  
  data <- candidates_to_check() %>% 
    mutate(new_label = input$new_label)
  
})

# write new name in db
observeEvent(input$new_label_submit, {
  
  this_candidate_id <- new_label() %>% pull(id)
  this_new_label <- new_label() %>% pull(new_label)
    
  query <- paste0("UPDATE ", db_tbl_er_candidates, " SET new_label = '", this_new_label, "' WHERE id =", this_candidate_id, ";")
  
  con <- connect_db()
  dbExecute(con, query)
  
  test_that(
    desc = "double-check if this_new_label was saved in db",
    expect_equal(
      tbl(con, db_tbl_er_candidates) %>%
        filter(id == this_candidate_id, !is.na(new_label)) %>%
        collect() %>% nrow(),
      1
    )
  )
  
  dbDisconnect(con)
  rm(con)
})

```


Column {data-width=70%}
-------------------------------------
```{r get_candidate_names}
candidate_names <- eventReactive(input$button_start, {
    candidates_to_check() %>% get_candidates_names()
  })

candidate_sources <- eventReactive(input$button_start, { 
    candidates_to_check() %>% select(source_1, source_2)
  })
```

### Candidate 1
    
```{r candiate_1_infos}

renderUI({
   tagList(
    h1(candidate_names() %>% pull(name_1)),
    p(paste("Source: ", candidate_sources() %>% pull(source_1))),
      p(paste("Item ID: ", candidate_names() %>% pull(id_1))),
    p(paste("Candidate Pair ID: ", candidate_names() %>% pull(id)))
   )
})

renderTable({
   if(candidate_sources() %>% pull(source_1) == "book_author") {
     inner_join(candidates_to_check() %>% distinct(source_1, id_1), 
           candidates$books_additional_infos %>% select(author_id, title = name, subtitle, publisher, isbn, location, year), 
           by = c("id_1" = "author_id")) %>% 
  select(-contains("source_"), -contains("id_")) 
    }
})


```
  

### Candidate 2
    
```{r candiate_2_infos}

renderUI({
   tagList(
    h1(candidate_names() %>% pull(name_2)),
    p(paste("Source: ", candidate_sources() %>% pull(source_2))),
    p(paste("Item ID: ", candidate_names() %>% pull(id_2))),
    p(paste("Candidate Pair ID: ", candidate_names() %>% pull(id)))
   )
})

renderTable({
   if(candidate_sources() %>% pull(source_2) == "book_author") {
     inner_join(candidates_to_check() %>% distinct(source_2, id_2), 
           candidates$books_additional_infos %>% select(author_id, title = name, subtitle, publisher, isbn, location, year), 
           by = c("id_2" = "author_id")) %>% 
  select(-contains("source_"), -contains("id_")) 
    }
})
```


```{r save_decision_after_click_in_db}

# when button click, then save decision in csv file
# it's an observeEvent, because nothing changes in app
# see: https://mastering-shiny.org/basic-reactivity.html#observers

# feedback https://mastering-shiny.org/action-feedback.html
# feedback: "Entscheidung gespeichert" can be just shown after save_decision()
# because if test in save_decision() fails, then whole app breaks :D

observeEvent(input$button_positive, {
  save_decision(candidates_to_check(), "positive", input$keep_candidate_2_label)
  # after decision for YES, set checkbox back to default state
  updateCheckboxInput(session,"keep_candidate_2_label", value = FALSE)
  showNotification("Entscheidung gespeichert. Klicke Start/Next...", duration = 2)
  })
  
observeEvent(input$button_negative, {
  save_decision(candidates_to_check(), "negative")
  showNotification("Entscheidung gespeichert. Klicke Start/Next...", duration = 2)
  })
  
observeEvent(input$button_not_sure, {
  save_decision(candidates_to_check(), "not_sure")
  showNotification("Entscheidung gespeichert. Klicke Start/Next...", duration = 2)
  })
```

Manual ER
=====================================

Row
-------------------------------------

### Add a manual entry for entity resolution

Du kannst hier manuell hinterlegen, wenn sich hinter zwei IDs die identische Identität verbirgt.

### Candidate 1
```{r manual_candidate_1}
fields <- c("manual_source_1", "manual_id_1", "manual_source_2", "manual_id_2")

selectInput("manual_source_1", "Quelle von Kandidat 1", choices = c("books" = "books", "poster" = "poster"))
textInput("manual_id_1", "ID von Kandidat 1", NA)
```

<!-- ### Info Candidate 1 -->

<!-- ```{r} -->

<!-- ``` -->

Row
-------------------------------------

### Save

```{r submit_manual_er}
actionButton("manual_submit", "Submit")
```

### Candidate 2
```{r manual_candidate_2}
selectInput("manual_source_2", "Quelle von Kandidat 2", choices = c("books" = "books", "poster" = "poster"))
textInput("manual_id_2", "ID von Kandidat 2", value = NA)
checkboxInput("manual_keep_candidate_2_label", "Behalte Name von Candidate 2")
```

<!-- ### Info Candidate 2 -->

<!-- ```{r} -->

<!-- ``` -->


```{r validate_save_manual_candidates}
# Validate input with ShinyValidator:
# Id MUST be given
# 1. Create an InputValidator object
iv_manual <- InputValidator$new()
# 2. Add validation rules
iv_manual$add_rule("manual_id_1", sv_required())
iv_manual$add_rule("manual_id_2", sv_required())
# 3. Start displaying errors in the UI
iv_manual$enable()
  
manual_input_data <- reactive({
     req(iv_manual$is_valid())
  
  data <- tibble(
    entities = "manual",
    source_1 = input$manual_source_1,
    source_2 = input$manual_source_2,
    id_1 = input$manual_id_1,
    id_2 = input$manual_id_2,
    decision = "positive"
  ) %>%
    mutate(keep_label_2 = ifelse(input$manual_keep_candidate_2_label == TRUE, 1, NA))
  data
})

# save data to db after click
observeEvent(input$manual_submit, {

  con <- connect_db()
  dbAppendTable(con, db_tbl_er_candidates, manual_input_data())

  test_that(
    desc = "double-check if this_decision was saved in db",
      expect_equal(
       tbl(con, db_tbl_er_candidates) %>%
           filter(entities == "manual", decision == "positive") %>%
           collect() %>%
           inner_join(manual_input_data()) %>%
         nrow(),
        1
      )
    )
      
  DBI::dbDisconnect(con)
  rm(con)
})
```

