---
title: "Compare Factgrid and Wikidata"
output: 
  flexdashboard::flex_dashboard:
    #orientation: rows
    vertical_layout: scroll
runtime: shiny
---


```{r setup, include=FALSE}
library(tidyverse)
library(glue)
library(SPARQL)
library(kabrutils)
library(here)
library(flexdashboard)
library(shinyFeedback)

source("functions.R", local = TRUE)
source("queries.R", local = TRUE)

shinyFeedback::useShinyFeedback()

```

Inputs {.sidebar}
=====================================

Which property should be compared?

```{r get_properties}
all_properties_raw <- get_properties()

all_properties <- all_properties_raw %>% 
  filter(str_detect(fg_property_type, "WikibaseItem|String|Time|Quantity|ExternalId|Url")) 
# not implemented: 
## CommonsMedia: URL must be cleaned

property_types_list <- all_properties %>% 
  select(fg_part_ofLabel, fg_part_of_id) %>% 
  distinct() %>% 
  arrange(fg_part_ofLabel) %>% 
  deframe()

```

```{r input_choose_properties_part_of, eval=FALSE}
# all_properties_list <- all_properties %>% 
#   select(fg_part_ofLabel, fg_part_of_id) %>% 
#   distinct() %>% 
#   arrange(fg_part_ofLabel) %>% 
#   deframe()
# 
# selectInput("property_parts_of", label = "Choose a topic:", all_properties_list, multiple = TRUE)
# actionButton("button_start_property_parts_of", "Query properties", class = "btn-primary")
# 
# observeEvent(input$button_start_property_parts_of, {
#   # first remove properties already chosen
#   removeUI(selector = "div:has(> #input_properties-label)")
#   # then insert chosen properties
#   insertUI(
#       selector = "#button_start_property_parts_of",
#       where = "afterEnd",
#       ui = checkboxGroupInput("input_properties", label = "Available properties:", choices = get_properties_list_part_of(all_properties, input$property_parts_of))
#     )
# })
```

```{r input_choose_properties_by_type}
all_properties_list <- all_properties %>% 
  count(fg_property_type, sort = T) %>%
  distinct(fg_property_type) %>% 
  arrange(fg_property_type) %>% 
  deframe()
names(all_properties_list) <- all_properties_list

selectInput("property_by_type", label = "Choose a property type:", all_properties_list, multiple = FALSE)
actionButton("button_start_property_by_type", "Query properties", class = "btn-primary")

observeEvent(input$button_start_property_by_type, {
  # first remove properties already chosen
  removeUI(selector = "div:has(> #input_properties-label)")
  # then insert chosen properties
  insertUI(
      selector = "#button_start_property_by_type",
      where = "afterEnd",
      ui = selectInput("input_properties", label = "Available properties:", choices = get_properties_list_by_type(all_properties, input$property_by_type))
        #checkboxGroupInput("input_properties", label = "Available properties:", choices = get_properties_list_by_type(all_properties, input$property_by_type))
    )
})
```

**Optional**: If you don't want to query all data in Factgrid, define a property-value-combination that needs to be satisfied. [Use this query to test it](https://database.factgrid.de/query/#%23defaultView%3ATable%0A%23%20Prefixes%20%20%20%20%0A%23%20Factgrid%0APREFIX%20fg%3A%20%3Chttps%3A%2F%2Fdatabase.factgrid.de%2Fentity%2F%3E%0APREFIX%20fgt%3A%20%3Chttps%3A%2F%2Fdatabase.factgrid.de%2Fprop%2Fdirect%2F%3E%0A%20%20%20%20%20%20%0A%23%20Wikidata%0APREFIX%20wdt%3A%20%3Chttp%3A%2F%2Fwww.wikidata.org%2Fprop%2Fdirect%2F%3E%0APREFIX%20wd%3A%20%3Chttp%3A%2F%2Fwww.wikidata.org%2Fentity%2F%3E%0A%23%20misc%0APREFIX%20wikibase%3A%20%3Chttp%3A%2F%2Fwikiba.se%2Fontology%23%3E%0A%20%20%20%20%20%20%0ASELECT%20DISTINCT%20%3Ffg_item%20%3Ffg_itemLabel%20where%20%7B%0A%20%20%20%20%20%20%0A%23%20labels%20from%20Factgrid%0A%20%20SERVICE%20wikibase%3Alabel%20%7B%20bd%3AserviceParam%20wikibase%3Alanguage%20%22%5BAUTO_LANGUAGE%5D%2Cen%22.%20%7D%20%20%0A%20%20%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%0A%20%20%23%20copy%20the%20next%20line%20in%20the%20filter%20field%3A%0A%20%20%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%23%0A%20%20%3Ffg_item%20fgt%3AP131%20fg%3AQ400012%20.%0A%7D). 

```{r input_filter_items}
textInput("item_filter_property", label = "Filter items: Property", placeholder = "P131")
textInput("item_filter_value", label = "Filter items: Value", placeholder = "Q400012")

actionButton("button_start_input_properties", "Submit", class = "btn-primary")
```

Please be patient, it may take a few moments if a lot of data is requested... If it takes too long, think about limiting the items that will be compared. Write a query for this in the field. 

```{r fetch_data}
comparison_raw <- eventReactive(input$button_start_input_properties, {
  
  if(is.null(input$input_properties)) {
    print(paste("chosen properties:", input$input_properties))
        shinyFeedback::showToast(
      "error", 
      "You have to select at least one property"
    )
  } else {
    
    #data <- get_comparison(input_properties() %>% select(fg_property), input$item_filter_property, input$item_filter_value)
     data <- get_comparison(input$input_properties, input$item_filter_property, input$item_filter_value)
    
    shinyFeedback::showToast(
      "success", 
      "data loaded"
    )
    
    data
  }
})
```

```{r get_chosen_properties}
input_properties <- eventReactive(input$button_start_input_properties, {
  data <- get_input_properties(all_properties, input$input_properties)
  cli::cli_alert_info("Input Properties dataframe:")
  print(data)
  data
})
```

Properties by Type
=====================================  

Properties you asked for:  {data-height=70, data-width=70}
-------------------------------------

### in both {.value-box}

```{r}
in_both_nrow <- reactive({
  nrow(in_both() %>% 
     distinct())
})

renderValueBox({
  valueBox(
    value = in_both_nrow(),
    icon = "fa-area-chart"
  )
})
```

### only in Factgrid {.value-box}

```{r}
only_in_fg_nrow <- reactive({
  nrow(only_in_fg() %>% 
     distinct())
})

renderValueBox({
  valueBox(
    value = only_in_fg_nrow(),
    icon = "fa-area-chart"
  )
})
```

### only in Wikidata {.value-box}

```{r}
only_in_wd_nrow <- reactive({
  nrow(only_in_wd() %>% 
         distinct(fg_item, wd_value_from_wd))
})

renderValueBox({
  valueBox(
    value = only_in_wd_nrow(),
    icon = "fa-area-chart"
  )
})
```


Row {.tabset .tabset-fade}
-------------------------------------
   
### Both, in Factgrid and Wikidata

#### Table
```{r filter_in_both}
in_both <- reactive({
  data <- comparison_raw() %>% 
    filter(is_same == "true") %>% 
    distinct(fg_item, fg_itemLabel, wd_item, fg_property_id, fg_value, fg_valueLabel, wd_value_from_fg, wd_value_from_wd) %>% 
    # add property info
    left_join(input_properties() %>% select(fg_property_id, fg_propertyLabel, fg_property, wd_property), by = "fg_property_id")
})
```

```{r show_table_in_both}
DT::renderDataTable(
  DT::datatable(

  in_both() %>% 
    left_join(input_properties() %>% distinct(fg_property_id, fg_property_type), by = "fg_property_id") %>% 
    # link for items
    create_link(url = fg_item, name = "Factgrid", new_col = fg_item_link) %>% 
    create_link(url = wd_item, name = "Wikidata", new_col = wd_item_link) %>% 
    mutate(item = paste0(fg_itemLabel, " (", fg_item_link, ", ", wd_item_link, ")")) %>% 
    # links for properties
    create_link(url = fg_property, name = "Factgrid", new_col = fg_property_link) %>% 
    create_link(url = wd_property, name = "Wikidata", new_col = wd_property_link) %>% 
    mutate(property = paste0(fg_propertyLabel, " (", fg_property_link, ", ", wd_property_link, ")")) %>% 
    
    # links for values
    create_link(url = fg_value, name = "Factgrid", new_col = fg_value_link) %>% 
    create_link(url = wd_value_from_wd, name = "Wikidata", new_col = wd_value_link) %>% 
    mutate(value = 
         case_when(
          fg_property_type == "WikibaseItem" ~ paste0(fg_valueLabel, " (", fg_value_link, ", ", wd_value_link, ")"),
          fg_property_type %in% c("Url", "CommonsMedia") ~ paste0('<a href = "', str_remove_all(wd_value_from_wd, "<|>"), '">', str_remove_all(wd_value_from_wd, "<|>"), '</a>'),
           TRUE ~ fg_valueLabel
           )
       ) %>% 
    
    # get relevant columns
    select(item, property, value) %>% 
     distinct(),
  escape = FALSE) # show link as link
)
```

### Only in Factgrid

```{r filter_only_in_fg}
only_in_fg <- reactive({
  
  data <- comparison_raw() %>% 
    filter(is.na(is_same), !is.na(fg_item)) %>% 
    distinct(fg_item, fg_itemLabel, wd_item, fg_property_id, fg_value, fg_valueLabel, wd_value_from_fg, wd_value_from_wd) %>% 
    # remove those that are present in wd
    anti_join(in_both(), by = c("fg_item", "wd_value_from_fg")) %>% 
    # add property info
    left_join(input_properties() %>% select(fg_property_id, fg_propertyLabel, fg_property, wd_property), by = "fg_property_id")
  
  })
```

#### Table
```{r show_table_only_in_fg}

# item var: fg_itemLabel
# value var: fg_valueLabel

DT::renderDataTable(
  DT::datatable(
  only_in_fg() %>% 
    left_join(input_properties() %>% distinct(fg_property_id, fg_property_type), by = "fg_property_id") %>% 
    # link for items
    create_link(url = fg_item, name = "Factgrid", new_col = fg_item_link) %>% 
    create_link(url = wd_item, name = "Wikidata", new_col = wd_item_link) %>% 
    mutate(item = paste0(fg_itemLabel, " (", fg_item_link, ", ", wd_item_link, ")")) %>% 
    # links for properties
    create_link(url = fg_property, name = "Factgrid", new_col = fg_property_link) %>% 
    create_link(url = wd_property, name = "Wikidata", new_col = wd_property_link) %>% 
    mutate(property = paste0(fg_propertyLabel, " (", fg_property_link, ", ", wd_property_link, ")")) %>% 
    
    # links for values
    create_link(url = fg_value, name = "Factgrid", new_col = fg_value_link) %>% 
    create_link(url = wd_value_from_fg, name = "Wikidata", new_col = wd_value_link) %>% 
    mutate(value = 
     case_when(
       fg_property_type == "WikibaseItem" ~ paste0(fg_valueLabel, " (", fg_value_link, ", ", wd_value_link, ")"),
       fg_property_type %in% c("Url", "CommonsMedia") ~ paste0('<a href = "', str_remove_all(fg_valueLabel, "<|>"), '">', str_remove_all(fg_valueLabel, "<|>"), '</a>'),
       TRUE ~ fg_valueLabel
       )
   ) %>% 
    
    # get relevant columns
    select(item, property, value) %>% 
     distinct(),
  escape = FALSE)
)

```

#### Import to Wikidata
This data can be imported to Wikidata. Click the button and paste it into [Wikidata's Quickstatements Tool](https://quickstatements.toolforge.org/#/).
```{r}
uiOutput("download_wikidata_import_UI")

import_in_wikidata <- reactive({
  data <- get_import_data(only_in_fg(), target = "wikidata", input_property_by_type = input$property_by_type) %>% 
     distinct()
  data
})

# create download button
output$download_wikidata_import_UI <- renderUI( {
  downloadButton("download_wikidata_import", "Download data for Wikidata import", class = "btn-primary", style = "width: 300px !important;")
})

# Add download handling
output$download_wikidata_import <- downloadHandler(
  filename = function() {
    "import_wikidata.csv"
  },
  content = function(file) {
    write.table(import_in_wikidata(), file, row.names = FALSE, quote = FALSE, sep = "\t")
  }
)
```

### Only in Wikidata

```{r filter_only_in_wd}
only_in_wd <- reactive({
  
  data <- comparison_raw() %>% 
    filter(is_same == "false") %>% 
    distinct(fg_item, fg_itemLabel, wd_item, fg_property_id, wd_value_from_wd, fg_value_from_wd) %>% 
    # remove those that are present in wd
    anti_join(in_both(), by = c("fg_item", "wd_value_from_wd")) %>%
    # add property info
    left_join(input_properties() %>% select(fg_property_id, fg_propertyLabel, fg_property, wd_property), by = "fg_property_id")
  
  })

```

#### Table

Labels from Wikidata are not shown (yet), because it would takes quite a time. 

```{r show_table_only_in_wd}
DT::renderDataTable(
  DT::datatable(
    only_in_wd() %>% 
      left_join(input_properties() %>% distinct(fg_property_id, fg_property_type), by = "fg_property_id") %>% 
     # link for items
      create_link(url = fg_item, name = "Factgrid", new_col = fg_item_link) %>% 
      create_link(url = wd_item, name = "Wikidata", new_col = wd_item_link) %>% 
      mutate(item = paste0(fg_itemLabel, " (", fg_item_link, ", ", wd_item_link, ")")) %>% 
      
      # links for properties
      create_link(url = fg_property, name = "Factgrid", new_col = fg_property_link) %>% 
      create_link(url = wd_property, name = "Wikidata", new_col = wd_property_link) %>% 
      mutate(property = paste0(fg_propertyLabel, " (", fg_property_link, ", ", wd_property_link, ")")) %>% 
      
      # links for values
      # no label, needs to be fetched from wikidata
      create_link(new_col = wd_value_link, url = wd_value_from_wd, name = "Wikidata") %>% 
      mutate(
        fg_value_from_wd_link = paste0('<a href="https://database.factgrid.de/wiki/Item:', fg_value_from_wd, '">Factgrid</a>'),
        value = 
          case_when(
            # for Wikibase Items
            !is.na(fg_value_from_wd) ~ paste0(" (", fg_value_from_wd_link, ", ", wd_value_link, ")"),
            fg_property_type == "WikibaseItem" & !is.na(fg_value_from_wd) ~ paste0(" (", fg_value_from_wd_link, ", ", wd_value_link, ")"),
            fg_property_type == "WikibaseItem" & is.na(fg_value_from_wd) ~ paste0(wd_value_link),
            # for Links
            fg_property_type %in% c("Url", "CommonsMedia") ~ paste0('<a href = "', str_remove_all(wd_value_from_wd, "<|>"), '">', str_remove_all(wd_value_from_wd, "<|>"), '</a>'),
            
            TRUE ~ as.character(wd_value_from_wd)
         )
    ) %>% 
      
      # get relevant columns
      select(item, property, value) %>% 
       distinct(),
  escape = FALSE)
)

```

#### Import to Factgrid

This data can be imported to Factgrid. Click the button and paste it into [Factgrid's Quickstatements Tool](https://database.factgrid.de/quickstatements/#/). Only those can be imported, that have a Factgrid-ID in Wikidata.

```{r import_to_factgrid}
uiOutput("download_factgrid_import_UI")

import_in_factgrid <- reactive({
  data <- get_import_data(only_in_wd(), target = "factgrid", input_property_by_type = input$property_by_type)# %>% 
    #distinct()
  print(data)
  data
})

# create download button
output$download_factgrid_import_UI <- renderUI( {
  downloadButton("download_factgrid_import", "Download data for Factgrid import", class = "btn-primary", style = "width: 300px !important;")
})

# Add download handling
output$download_factgrid_import <- downloadHandler(
  filename = function() {
    "import_factgrid.csv"
  },
  content = function(file) {
    write.table(import_in_factgrid(), file, row.names = FALSE, quote = FALSE, sep = "\t")
  }
)
```