---
title: "Compare Factgrid and Wikidata"
output: 
  flexdashboard::flex_dashboard:
    #orientation: rows
    vertical_layout: scroll
runtime: shiny
---


```{r setup, include=FALSE}
library(tidyverse)
library(glue)
library(SPARQL)
library(kabrutils)
library(here)
library(flexdashboard)
library(WikidataR)
library(clipr)
config_file <- "data-publishing/factgrid/config.yml"
config <- yaml::read_yaml(here(config_file))
```


```{r fetch_available_properties}

#' Get properties related to persons
#' @return dataframe
get_person_properties <- function() {
  tmp <- readr::read_file("https://raw.githubusercontent.com/cutterkom/remove-na-lgbtiq-queer-knowledge-graph/main/data-publishing/factgrid/queries/get_all_properties_person_with_corresponding_prop.rq") %>% 
  sparql_to_tibble(endpoint = config$connection$sparql_endpoint) %>% 
  mutate(
    fg_property_id = str_extract(fg_property, "P[0-9]+"),
    wd_property_id = str_extract(wd_property, "P[0-9]+"),
    label = paste0(fg_propertyLabel, " (FG: ", fg_property_id, ", WD: ", wd_property_id, ")"),
    label = str_remove_all(label, '"|@en')
    ) 
    
  return(tmp)
}
```

Refering to Persons
=====================================  
 {.sidebar}
-------------------------------------

Which property should be compared?

```{r get_properties}
persons_properties <- get_person_properties() %>% arrange(fg_propertyLabel)

persons_properties_list  <- persons_properties %>% 
  select(label, fg_property_id) %>% 
  deframe()

checkboxGroupInput("person_properties", label = "Properties available that refer to persons", choices = persons_properties_list)

actionButton("button_start_person_properties", "Submit", class = "btn-primary")
```

Please be patient, it may take a few moments if a lot of data is requested... 

```{r get_comparison_fun}

#' Get comparison data based on property
#' 
#' This functions queries data on Factgrid looking for properties related to persons that have a corresponding Wikidata property. Utilizes federated queries.
#' @param input chosen properties in input$person_properties
#' @return dataframe
get_comparison <- function(input) {
  persons_properties %>% 
    filter(fg_property_id %in% input) %>% 
    pmap_dfr(function(...) {
      current <- tibble(...)
      cli::cli_alert_info("Fetch data for: {current$fg_propertyLabel}")
      fg_property_id <- current$fg_property_id
      
      query <- paste0( '
      #defaultView:Table
      
      # Prefix standard 
      # Factgrid
      PREFIX fg: <https://database.factgrid.de/entity/>
      PREFIX fgt: <https://database.factgrid.de/prop/direct/>
      
      # Wikidata
      PREFIX wdt: <http://www.wikidata.org/prop/direct/>
      PREFIX wd: <http://www.wikidata.org/entity/>
      # misc
      PREFIX wikibase: <http://wikiba.se/ontology#>
      PREFIX bd: <http://www.bigdata.com/rdf#>
      PREFIX schema: <http://schema.org/>
      
      SELECT DISTINCT ?fg_item ?fg_itemLabel ?wd_item ?fg_property ?fg_propertyLabel ?wd_property ?fg_value ?fg_valueLabel ?wd_value_from_fg ?wd_value_from_wd ?fg_value_from_wd ?is_same where {
      
        # labels from Factgrid
        SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". }  
      
        # starting points
        # which property in focus?
        # todo in one var
        BIND(fgt:', fg_property_id, ' as ?fg_property)
        BIND(fg:', fg_property_id, ' as ?fg_property_as_item)
        # set property of corresponding wikidata property; is constant
        BIND(fgt:P343 as ?fg_corr_wd)
      
        # transform wikidata qid in factgrid to wikidata entity iri
        ?link schema:about ?fg_item .
        ?link schema:isPartOf <https://www.wikidata.org/> . 
        ?link schema:name ?qid.
        BIND(IRI(CONCAT(STR(wd:), ?qid)) AS ?wd_item)
      
        # get corresponding wikidata property
        # and store as IRI
        ?fg_property_as_item ?fg_corr_wd ?fg_corr_wd_value .
        BIND(IRI(CONCAT(STR(wdt:), ?fg_corr_wd_value)) AS ?wd_property)
      
        # get factgrid statement
        ?fg_item ?fg_property ?fg_value .
        # get wikidata qid of fg_value
        # used later for comparison
        ?link_value schema:about ?fg_value .
        ?link_value schema:isPartOf <https://www.wikidata.org/> . 
        ?link_value schema:name ?qid_value.
        BIND(IRI(CONCAT(STR(wd:), ?qid_value)) AS ?wd_value_from_fg)
      
        # now change to wikidata as data source
        # get value from wikidata
        SERVICE <https://query.wikidata.org/sparql> {
            OPTIONAL {
                ?wd_item ?wd_property ?wd_value_from_wd.
            }

        }
        BIND (IF(?wd_value_from_fg = ?wd_value_from_wd, "true", "false") AS ?is_same)
      }
      
      ')
      
      res <- sparql_to_tibble(query, endpoint = config$connection$sparql_endpoint) 
      print(res)
       if(nrow(res) > 0) {
        res <- res %>% 
          mutate(
            fg_property_id = str_extract(fg_property, "P[0-9]+"),
            wd_property_id = str_extract(wd_property, "P[0-9]+"))
       } else {
        res <- tibble(
          fg_item = NA_character_,
          fg_itemLabel = NA_character_,
          wd_item = NA_character_,
          fg_property = NA_character_,
          fg_propertyLabel = NA_character_,
          wd_property = NA_character_,
          fg_value = NA_character_,
          fg_valueLabel = NA_character_,
          wd_value_from_fg = NA_character_,
          wd_value_from_wd = NA_character_,
          is_same = NA_character_,
          fg_property_id = NA_character_,
          wd_property_id = NA_character_
        )
       }
      cli::cli_alert_success("data fetched")
     res
    })
}

```

```{r create_link_fun}

#' Function that creates a link from href and string
#' @param data dataframe with cols
#' @param url href for link
#' @param name string for inside <a href>name</a>
#' @param new_col name of new col
create_link <- function(data, url, name, new_col) {
  data %>% 
    mutate(
      {{new_col}} := str_replace({{url}}, "<", "<a href='"),
      {{new_col}} := str_replace({{new_col}}, ">", "' target='_blank'>"),
      {{new_col}} := paste0({{new_col}}, {{name}}, "</a>")
    )
}
```

```{r get_import_data}
get_import_data <- function(data, target) {
  
  tmp <- comparison_raw() %>% 
    inner_join(data %>% distinct(fg_item, wd_value_from_fg), by = c("fg_item", "wd_value_from_fg")) %>% 
    mutate(
      fg_item_id = str_extract(fg_item, "Q[0-9]+"),
      fg_value_id = str_extract(fg_value, "Q[0-9]+"),
      wd_item_id = str_extract(wd_item, "Q[0-9]+"),
      wd_value_from_fg_id = str_extract(wd_value_from_fg, "Q[0-9]+")
    )
  
  if (target == "wikidata") {
    import_data <- tmp %>% 
      mutate(fg_value_id = paste0('"', fg_value_id, '"')) %>% 
      # add source of Factgrid-Object
      bind_cols(tibble(source = "S8168", time = "S813", timestamp = paste0("+", Sys.Date(), "T00:00:00Z/", 11))) %>% 
      select(item = wd_item_id, property = wd_property_id, value = wd_value_from_fg_id, 
             source, source_value = fg_value_id, 
             time, timestamp) %>% 
      distinct()
  } else if (target == "factgrid") {
    # import_data <- tmp %>% 
    #   mutate(wd_item_id = str_extract(wd_item, "Q[0-9]+"))
    #   bind_cols(tibble(source = "S8168", time = "S813", timestamp = paste0("+", Sys.Date(), "T00:00:00Z/", 11))) %>% 
    #   select(item = wd_item_id, property = wd_property_id, value = wd_value_from_fg_id, source, source_value = fg_item_id, time, timestamp) %>% 
    #   distinct()
  } else {
    stop("no valid target. Only `wikidata` or `factgrid` allowed")
  }
  
}
```


```{r fetch_data}
comparison_raw <- eventReactive(input$button_start_person_properties, {
    data <- get_comparison(input$person_properties)
    data
  })
```

Properties you asked for:  {data-height=70}
-------------------------------------

### in both {.value-box}

```{r}
in_both_nrow <- reactive({
  nrow(in_both() %>% 
     distinct())
})

renderValueBox({
  valueBox(
    value = in_both_nrow(),
    icon = "fa-area-chart"
  )
})
```

### only in Factgrid {.value-box}

```{r}
only_in_fg_nrow <- reactive({
  nrow(only_in_fg() %>% 
     distinct())
})

renderValueBox({
  valueBox(
    value = only_in_fg_nrow(),
    icon = "fa-area-chart"
  )
})
```

### only in Wikidata {.value-box}

```{r}
only_in_wd_nrow <- reactive({
  nrow(only_in_wd() %>% 
         distinct(fg_item, wd_value_from_wd))
})

renderValueBox({
  valueBox(
    value = only_in_wd_nrow(),
    icon = "fa-area-chart"
  )
})
```


Row {.tabset .tabset-fade}
-------------------------------------
   
### Both, in Factgrid and Wikidata

#### Table
```{r filter_in_both}
in_both <- reactive({
  data <- comparison_raw() %>% 
    filter(is_same == "true") %>% 
    distinct(fg_item, fg_itemLabel, wd_item, fg_property_id, fg_value, fg_valueLabel, wd_value_from_fg, wd_value_from_wd) %>% 
    # add property info
    left_join(persons_properties %>% select(fg_property_id, fg_propertyLabel, fg_property, wd_property), by = "fg_property_id")
})

```

```{r show_table_in_both}
DT::renderDataTable(
  DT::datatable(

  in_both() %>% 
    # link for items
    create_link(url = fg_item, name = "Factgrid", new_col = fg_item_link) %>% 
    create_link(url = wd_item, name = "Wikidata", new_col = wd_item_link) %>% 
    mutate(item = paste0(fg_itemLabel, " (", fg_item_link, ", ", wd_item_link, ")")) %>% 
    # links for properties
    create_link(url = fg_property, name = "Factgrid", new_col = fg_property_link) %>% 
    create_link(url = wd_property, name = "Wikidata", new_col = wd_property_link) %>% 
    mutate(property = paste0(fg_propertyLabel, " (", fg_property_link, ", ", wd_property_link, ")")) %>% 
    
    # links for values
    create_link(url = fg_value, name = "Factgrid", new_col = fg_value_link) %>% 
    create_link(url = wd_value_from_fg, name = "Wikidata", new_col = wd_value_link) %>% 
    mutate(value = paste0(fg_valueLabel, " (", fg_value_link, ", ", wd_value_link, ")")) %>% 
    
    # get relevant columns
    select(item, property, value) %>% 
     distinct(),
  escape = FALSE) # show link as link
)


```

### Only in Factgrid

```{r filter_only_in_fg}
only_in_fg <-  reactive({
  
  data <- comparison_raw() %>% 
    filter(is.na(is_same), !is.na(fg_item)) %>% 
    distinct(fg_item, fg_itemLabel, wd_item, fg_property_id, fg_value, fg_valueLabel, wd_value_from_fg, wd_value_from_wd) %>% 
    # remove those that are present in wd
    anti_join(in_both(), by = c("fg_item", "wd_value_from_fg")) %>% 
    # add property info
    left_join(persons_properties %>% select(fg_property_id, fg_propertyLabel, fg_property, wd_property), by = "fg_property_id")
  
  })
```

#### Table
```{r show_table_only_in_fg}
DT::renderDataTable(
  DT::datatable(
  only_in_fg() %>% 
    # link for items
    create_link(url = fg_item, name = "Factgrid", new_col = fg_item_link) %>% 
    create_link(url = wd_item, name = "Wikidata", new_col = wd_item_link) %>% 
    mutate(item = paste0(fg_itemLabel, " (", fg_item_link, ", ", wd_item_link, ")")) %>% 
    # links for properties
    create_link(url = fg_property, name = "Factgrid", new_col = fg_property_link) %>% 
    create_link(url = wd_property, name = "Wikidata", new_col = wd_property_link) %>% 
    mutate(property = paste0(fg_propertyLabel, " (", fg_property_link, ", ", wd_property_link, ")")) %>% 
    
    # links for values
    create_link(url = fg_value, name = "Factgrid", new_col = fg_value_link) %>% 
    create_link(url = wd_value_from_fg, name = "Wikidata", new_col = wd_value_link) %>% 
    mutate(value = paste0(fg_valueLabel, " (", fg_value_link, ", ", wd_value_link, ")")) %>% 
    
    # get relevant columns
    select(item, property, value) %>% 
     distinct(),
  escape = FALSE)
)

```

#### Import to Wikidata
This data can be imported to Wikidata. Click the button and paste it into [Wikidata's Quickstatements Tool](https://quickstatements.toolforge.org/#/).
```{r}
actionButton("button_import_to_wikidata", "Copy Data", class = "btn-primary")
import_in_wikidata <- reactive({
  data <- get_import_data(only_in_fg(), target = "wikidata") %>% 
     distinct()
})

observeEvent(input$button_import_to_wikidata, {
    clipr::write_clip(import_in_wikidata(), allow_non_interactive = TRUE)
})
```

### Only in Wikidata

```{r filter_only_in_wd}
only_in_wd <- reactive({
  
  data <- comparison_raw() %>% 
    filter(is_same == "false") %>% 
    distinct(fg_item, fg_itemLabel, wd_item, fg_property_id, fg_value, fg_valueLabel, wd_value_from_fg, wd_value_from_wd) %>% 
    # remove those that are present in wd
    anti_join(in_both(), by = c("fg_item", "wd_value_from_wd")) %>%
    # add property info
    left_join(persons_properties %>% select(fg_property_id, fg_propertyLabel, fg_property, wd_property), by = "fg_property_id")
  
  })

```

#### Table

Labels from Wikidata are not shown (yet), because it would takes quite a time. 
```{r show_table_only_in_wd}
DT::renderDataTable(
  DT::datatable(
    only_in_wd() %>% 
     # link for items
      create_link(url = fg_item, name = "Factgrid", new_col = fg_item_link) %>% 
      create_link(url = wd_item, name = "Wikidata", new_col = wd_item_link) %>% 
      mutate(item = paste0(fg_itemLabel, " (", fg_item_link, ", ", wd_item_link, ")")) %>% 
      
      # links for properties
      create_link(url = fg_property, name = "Factgrid", new_col = fg_property_link) %>% 
      create_link(url = wd_property, name = "Wikidata", new_col = wd_property_link) %>% 
      mutate(property = paste0(fg_propertyLabel, " (", fg_property_link, ", ", wd_property_link, ")")) %>% 
      
      # links for values
      # no label, needs to be fetched from wikidata
      create_link(new_col = wd_value_link, url = wd_value_from_wd, name = "Wikidata") %>% 
      mutate(value = paste0(wd_value_link)) %>% 
      
      # get relevant columns
      select(item, property, value) %>% 
       distinct(),
  escape = FALSE)
)

```


```{r eval=F}
#### Import to Factgrid
This data can be imported to Factgrid. Click the button and paste it into [Factgrid's Quickstatements Tool](https://database.factgrid.de/quickstatements/#/).
actionButton("button_import_to_factgrid", "Copy Data", class = "btn-primary")
import_in_factgrid <- reactive({
  data <- get_import_data(only_in_wd(), target = "factgrid") %>% 
    distinct()
})

observeEvent(input$button_import_to_factgrid, {
    clipr::write_clip(import_in_factgrid(), allow_non_interactive = TRUE)
})
```