---
title: "Stock Taking: How big is the queer Wikidata"
date: "2022-03-18T00:00:00"
status: "private"
slug: "stock-taking-how-big-is-queer-wikidata"
output: rmdformats::robobook
---

My reference for how much queer^[I use queer, LGBT and LGBTIQ* as synonyms.] history is in linked data is Wikidata. 
So the first step is to take stock.

I do that in two ways: 

1. Take **LGBTIQ* properties** and fetch all items that have this relation
2. Take **main LGBTIQ* items** and look up which items have a relation

## Setup

Of course, I could run many different SPARQL queries.
A more efficient way for my workflows is using the package {tidywikidatar} that offers, like the name suggests, tools to retrieve data in a [tidy (read: recangular) way](https://r4ds.had.co.nz/tidy-data.html).
But I'll also use SPARQL for a bit more complex queries with {WikidataR}.

```{r, message=F, warning=F}
library(tidywikidatar) # tidy queries
library(WikidataR) # SPARQL queries
library(tidyverse)

# use caching from tidywikidatar
tw_enable_cache()
tw_set_cache_folder(path = fs::path(fs::path_home_r(), "R", "tw_data"))
tw_set_language(language = "en")
tw_create_cache_folder(ask = FALSE)
tw_get_cache_folder()
```

## LGBTIQ* main items

The [Wikidata:WikiProject LGBT](https://www.wikidata.org/wiki/Wikidata:WikiProject_LGBT/Models) lists main items and suggested properties for these items.
I transferred these to a table in order to use them to get data from Wikidata in bulk.

```{r, warning=F, message=F}
lgbt_items <- read_tsv("../../../static/wikidata-queer-items.tsv") %>% 
  mutate(q = str_extract(item, "Q[0-9]+"),
         p = str_extract_all(related_property, "P[0-9]+")) %>% 
  unnest(p)
```

I take every item-property-combination from the input file and get all sub-items.

```{r, warning=FALSE, message=FALSE, cache=TRUE}
subitems <- lgbt_items %>%
  pmap_dfr(function(...) {
    current <- tibble(...) %>% 
      select(item, topic, q, p)

    if (current$topic %in% c("Places & organizations")) {
      current %>%
        tw_query() %>%
        bind_cols(current) %>% 
        mutate(country_id = tw_get_p(id, "P17")) %>% 
        unnest(country_id) %>% 
        mutate(country = tw_get_label(id = country_id))
    } else {
      current %>%
        tw_query() %>%
        bind_cols(current)
    }
  }) %>% 
  select(topic, item, p, p, subitem = id, label, description, country_id, country)

DT::datatable(subitems)
```

So there are `r nrow(subitems)` I can find with this approach.
But be aware, that there are duplicates, e.g. because I ask partly for two properties or one subitem can be an LGBT bar as well as a lesbian bar.

When I take this into account, there are `r subitems %>% distinct(subitem, label, topic) %>% nrow()` distinct subitems available. 

Counted by topic:

```{r}
topic_share <- subitems %>%
  count(topic) %>%
  mutate(pct = n / sum(n)) %>% 
  arrange(desc(pct))

DT::datatable(topic_share)
```

As a chart: 

```{r}
topic_share %>% 
  ggplot(aes(reorder(topic, -n), n)) +
  geom_col() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90))
  
```

Counted by topic and item:

```{r}
topic_item_share <- subitems %>%
  count(topic, item) %>%
  mutate(pct = n / sum(n)) %>% 
  arrange(desc(pct))

DT::datatable(topic_item_share)
```

That's quite a long longtail:

```{r}
topic_item_share %>% 
  ggplot(aes(reorder(item, -n), n)) +
  geom_col() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90))
```


Since my project revolves around German LGBTIQ* history, country information is important.
For items having the topic `Places & organizations` is queried the country information if available.

Just 16 location that are LGBTIQ* related are based in Germany.
Or are labeled as some kind of queer location. For example the famous hotel, bar, sauna "Deutsche Eiche" in Munich ([German Wikipedia](https://de.wikipedia.org/wiki/Hotel_Deutsche_Eiche)) has no indication of its role for the gay scene on [Wikidata](https://www.wikidata.org/wiki/Q1630878).

```{r}
subitems %>% 
  filter(topic == "Places & organizations") %>% 
  count(country, sort = T)
```

Also, no LGBT archives from Germany can be found in the data, even if they exist as Wikipedia item, for instance the archive I am with: [Forum Queeres Archiv](https://www.wikidata.org/wiki/Q100743737). There are really some low hanging fruits out there...

## LGBTIQ* properties

Now the same with properties. 

```{r, warning=F, message=F}
lgbt_properties <- read_tsv("../../../static/wikidata-queer-properties.tsv") %>% 
  rename(p = id)

DT::datatable(lgbt_properties)
```

```{r}
items_with_properties <- lgbt_properties %>%
  filter(!data_type %in% c("Item", "Lexeme")) %>% 
  pmap_dfr(function(...) {
    current <- tibble(...) %>% 
      select(property, p)
    print(current)
    current %>% pull(p) %>% 
      tw_get_all_with_p() %>% 
      bind_cols(current)
  }) %>% 
  select(property, p, everything())

DT::datatable(items_with_properties)
```

As counts by propery: 

```{r}
properties_share <- items_with_properties %>%
  count(property) %>%
  mutate(pct = n / sum(n)) %>% 
  arrange(desc(pct))

DT::datatable(properties_share)
```

## Persons, that are not heterosexual

This time, I use plain SPARQL to get the data, [based on that query](https://www.wikidata.org/wiki/Wikidata:WikiProject_LGBT/Queries#LGBT_people_born,_residing,_or_working_in_New_York_City).

```{r}
non_heterosexuals <- query_wikidata(
  '
  SELECT DISTINCT ?person ?personLabel ?sexualorientationLabel ?sexorgenderLabel
  WHERE {
    ?person wdt:P31 wd:Q5 . #?person is a human
  
    { 
      ?person wdt:P21 ?sexorgender. #?person has ?sexorgender
      #?sexorgender is not male, female, cisgender male, cigender female, or cisgender person
      FILTER(?sexorgender NOT IN (wd:Q6581097, wd:Q6581072, wd:Q15145778, wd:Q15145779, wd:Q1093205)). 
    } UNION {
      ?person wdt:P91 ?sexualorientation . #?person has ?sexualorientation
      FILTER(?sexualorientation != wd:Q1035954). #?sexualorientation is not heterosexual
    }
    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". }
}'
) %>% as_tibble()

DT::datatable(non_heterosexuals)
```

```{r}
non_heterosexuals %>% 
  count(sexualorientationLabel, sort = T) %>% 
  DT::datatable()
```
Woah, what are those `r non_heterosexuals %>% filter(is.na(sexualorientationLabel)) %>% nrow()` that have an empty label?
Are these subversive attempts to undermine labels that deal with sexual orientation?

A sample shows: Not impossible, but probably not.
For instance, quite a lot are painting with humans. Or "anonymous masters", artists whose sex or gender is unknown, might be queer persons, might be not.
So this must be taken with a grain of salt.

(For quite some time now, I want to write a blog post about [the book "Queer Data" by Kevin Guyan](https://kevinguyan.com/queer-data/). This would be an excellent link.)

Since the focus of Remove NA is the city of Munich I want to see: What are non-heterosexual or non-cis persons that are somehow linked to Munich? 

```{r}
munich <- query_wikidata('
# LGBT that are linked to Munich
SELECT DISTINCT ?person ?personLabel ?sexualorientationLabel ?sexorgenderLabel
  WHERE {
    ?person wdt:P31 wd:Q5 . #?person is a human
    BIND(wd:Q1726 as ?place). #Q1726 is Munich
    
    { 
      ?person wdt:P21 ?sexorgender. #?person has ?sexorgender
      #?sexorgender is not male, female, cisgender male, cigender female, or cisgender person
      FILTER(?sexorgender NOT IN (wd:Q6581097, wd:Q6581072, wd:Q15145778, wd:Q15145779, wd:Q1093205)). 
    } UNION {
      ?person wdt:P91 ?sexualorientation . #?person has ?sexualorientation
      FILTER(?sexualorientation != wd:Q1035954). #?sexualorientation is not heterosexual
    }
    
    {
      ?person wdt:P19/wdt:P131* ?place. #?person was born in ?place
    }
    UNION {
      ?person wdt:P551/wdt:P131* ?place. #?person resides in ?place
    }
    UNION {
      ?person wdt:P937/wdt:P131* ?place. #?person works in ?place
    }

    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". }
}
') %>% as_tibble()

DT::datatable(munich)
```

## What comes next

These queries are just a first glimpse into queer Wikidata and what I want to do in my project `Remove NA`.